---
title: "Hands-On Exercise 1: Grammar of Graphics with ggplot2"
author: "Zhang Xuerong"
format: html
editor: visual
---

```{r}
# Load required packages
install.packages("pacman", repos = "https://cloud.r-project.org/")
pacman::p_load(tidyverse)
```

```{r}
# Load the data
#install.packages("readr")

library(readr)

exam_data <- read_csv("hands-on_exercise1/data/Exam_data.csv")
```

## R Graphics VS ggplot

```{r}
#install.packages("ggplot2")  # run only if not installed
library(ggplot2)

ggplot(data=exam_data, aes(x = MATHS)) +
  geom_histogram(bins=10, 
                 boundary = 100,
                 color="black", 
                 fill="grey") +
  ggtitle("Distribution of Maths scores")
```


```{r}
hist(exam_data$MATHS)
```

## Geometric Objects

### Bar Chart

```{r}
ggplot(data = exam_data, aes(x = RACE)) + 
  geom_bar()
```

### Dot Plot

```{r}
ggplot(data = exam_data, aes(x = MATHS)) +
  geom_dotplot(binwidth = 2.5, dotsize = 0.5) +
  scale_y_continuous(NULL, breaks = NULL)
```

### Histogram

```{r}
ggplot(data = exam_data, aes(x = MATHS)) +
  geom_histogram()
```

```{r}
ggplot(data = exam_data, aes(x = MATHS)) +
  geom_histogram(bins = 20, fill = "light blue", color = "black")
```

### Histogram by Gender

```{r}
ggplot(data = exam_data, aes(x = MATHS, fill = GENDER)) +
  geom_histogram(bins = 20, color = "grey30")
```

### Density Plot

```{r}
ggplot(data = exam_data, aes(x = MATHS)) +
  geom_density()
```

```{r}
ggplot(data = exam_data, aes(x = MATHS, colour = GENDER)) +
  geom_density()
```

### Boxplot

```{r}
ggplot(data = exam_data, aes(x = GENDER, y = MATHS)) +
  geom_boxplot()
```

```{r}
ggplot(data = exam_data, aes(x = GENDER, y = MATHS)) +
  geom_boxplot(notch = TRUE)
```

### Violin Plot

```{r}
ggplot(data = exam_data, aes(x = GENDER, y = MATHS)) +
  geom_violin()
```

### Scatterplot

```{r}
ggplot(data = exam_data, aes(x = MATHS, y = ENGLISH)) +
  geom_point()
```

### Boxplot + Points

```{r}
ggplot(data = exam_data, aes(x = GENDER, y = MATHS)) +
  geom_boxplot() +
  geom_point(position = "jitter", size = 0.5)
```

### Boxplot with Mean (stat_summary)

```{r}
ggplot(data = exam_data, aes(x = GENDER, y = MATHS)) +
  geom_boxplot() +
  stat_summary(fun = "mean", geom = "point", colour = "red", size = 4)
```

### Boxplot with Mean (geom_point)

```{r}
ggplot(data = exam_data, aes(x = GENDER, y = MATHS)) +
  geom_boxplot() +
  geom_point(stat = "summary", fun = "mean", colour = "red", size = 4)
```

### Scatterplot with Smooth Line

```{r}
ggplot(data = exam_data, aes(x = MATHS, y = ENGLISH)) +
  geom_point() +
  geom_smooth(size = 0.5)
```

```{r}
ggplot(data = exam_data, aes(x = MATHS, y = ENGLISH)) +
  geom_point() +
  geom_smooth(method = lm, linewidth = 0.5)
```

## Facets

### facet_wrap()

```{r}
ggplot(data = exam_data, aes(x = MATHS)) +
  geom_histogram(bins = 20) +
  facet_wrap(~CLASS)
```

### facet_grid()

```{r}
ggplot(data = exam_data, aes(x = MATHS)) +
  geom_histogram(bins = 20) +
  facet_grid(~CLASS)
```

## Coordinate Systems

### Horizontal Bar Chart

```{r}
ggplot(data = exam_data, aes(x = RACE)) +
  geom_bar() +
  coord_flip()
```

### Fix Axis Limits

```{r}
ggplot(data = exam_data, aes(x = MATHS, y = ENGLISH)) +
  geom_point() +
  geom_smooth(method = lm, size = 0.5) +
  coord_cartesian(xlim = c(0, 100), ylim = c(0, 100))
```

## Themes

### theme_gray (default)

```{r}
ggplot(data = exam_data, aes(x = RACE)) +
  geom_bar() +
  coord_flip() +
  theme_gray()
```

### theme_classic

```{r}
ggplot(data = exam_data, aes(x = RACE)) +
  geom_bar() +
  coord_flip() +
  theme_classic()
```

### theme_minimal

```{r}
ggplot(data = exam_data, aes(x = RACE)) +
  geom_bar() +
  coord_flip() +
  theme_minimal()
```
