---
title: "Hands-on Exercise 4: Visual Statistical Analysis"
format: html
editor: visual
---

```{r setup, include=FALSE}
# Load necessary packages
pacman::p_load(ggstatsplot, tidyverse)
```

## 1. Importing Data

```{r}
# Read in the Exam.csv file
exam <- read_csv("/Users/sharon/OneDrive - Singapore Management University/isss608data/hands-on_exercise2/Exam_data.csv")

# View the structure of the dataset
glimpse(exam)
```

## 2. One-Sample Test on English Scores

```{r}
# Visualizing a Bayesian one-sample test for English scores
set.seed(1234)

gghistostats(
  data = exam,
  x = ENGLISH,
  type = "bayes",
  test.value = 60,
  xlab = "English scores"
)
```

## 3. Interpreting Bayes Factor

The Bayes Factor (BF10) quantifies the evidence for the alternative hypothesis (H1) compared to the null hypothesis (H0). Interpretation based on Jeffreys’ scale:

-   **BF \< 1**: Evidence for H0
-   **1 \< BF \< 3**: Anecdotal evidence for H1
-   **3 \< BF \< 10**: Moderate evidence for H1
-   **BF \> 10**: Strong evidence for H1

## 4. Two-Sample Mean Test: Maths Scores by Gender

```{r}
# Visualizing a non-parametric test for Maths scores by Gender
ggbetweenstats(
  data = exam,
  x = GENDER, 
  y = MATHS,
  type = "np",
  messages = FALSE
)
```

## 5. One-Way ANOVA Test: English Scores by Race

```{r}
# One-way ANOVA with post-hoc pairwise comparisons for English scores by Race
ggbetweenstats(
  data = exam,
  x = RACE, 
  y = ENGLISH,
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "s",
  p.adjust.method = "fdr",
  messages = FALSE
)
```

## 6. Correlation Test: Maths vs English Scores

```{r}
# Visual test of correlation between Maths and English scores
ggscatterstats(
  data = exam,
  x = MATHS,
  y = ENGLISH,
  marginal = FALSE
)
```

## 7. Test of Association: Binned Maths Scores vs Gender

```{r}
# Create bins for Maths scores
exam1 <- exam %>% 
  mutate(MATHS_bins = cut(MATHS, breaks = c(0, 60, 75, 85, 100)))

# Association test between binned Maths scores and Gender
ggbarstats(
  data = exam1, 
  x = MATHS_bins, 
  y = GENDER
)
```

---
