---
title: "Hands-on Exercise 2: Beyond ggplot2 Fundamentals"
format: html
editor: visual
---

## Setup

```{r}
# Install and load necessary packages
pacman::p_load(
  tidyverse, ggrepel, ggthemes, hrbrthemes, patchwork
)
```

## Import Data

```{r}
# Read the exam data CSV
exam_data <- read_csv("/Users/sharon/OneDrive - Singapore Management University/isss608data/hands-on_exercise2/Exam_data.csv")
```

## ggrepel: Prevent Overlapping Labels

```{r}
ggplot(data = exam_data, aes(x = MATHS, y = ENGLISH)) +
  geom_point() +
  geom_smooth(method = lm, size = 0.5) +
  geom_label_repel(aes(label = ID), fontface = "bold") +
  coord_cartesian(xlim = c(0, 100), ylim = c(0, 100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```

## ggthemes: Using Predefined Themes

### Base ggplot2 Theme

```{r}
ggplot(data = exam_data, aes(x = MATHS)) +
  geom_histogram(bins = 20, boundary = 100, color = "grey25", fill = "grey90") +
  theme_gray() +
  ggtitle("Distribution of Maths scores")
```

### Economist Theme

```{r}
ggplot(data = exam_data, aes(x = MATHS)) +
  geom_histogram(bins = 20, boundary = 100, color = "grey25", fill = "grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_economist()
```

## hrbrthemes: Typography-Centric Themes

### Default hrbrthemes

```{r}
ggplot(data = exam_data, aes(x = MATHS)) +
  geom_histogram(bins = 20, boundary = 100, color = "grey25", fill = "grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_ipsum()
```

### Customized Typography and Grid

```{r}
ggplot(data = exam_data, aes(x = MATHS)) +
  geom_histogram(bins = 20, boundary = 100, color = "grey25", fill = "grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_ipsum(axis_title_size = 18, base_size = 15, grid = "Y")
```

## Composite Figures with Patchwork

### Create Individual Plots

```{r}
p1 <- ggplot(data = exam_data, aes(x = MATHS)) +
  geom_histogram(bins = 20, boundary = 100, color = "grey25", fill = "grey90") +
  coord_cartesian(xlim = c(0, 100)) +
  ggtitle("Distribution of Maths scores")

p2 <- ggplot(data = exam_data, aes(x = ENGLISH)) +
  geom_histogram(bins = 20, boundary = 100, color = "grey25", fill = "grey90") +
  coord_cartesian(xlim = c(0, 100)) +
  ggtitle("Distribution of English scores")

p3 <- ggplot(data = exam_data, aes(x = MATHS, y = ENGLISH)) +
  geom_point() +
  geom_smooth(method = lm, size = 0.5) +
  coord_cartesian(xlim = c(0, 100), ylim = c(0, 100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```

### Combine Two Plots Horizontally

```{r}
p1 + p2
```

### Combine Three Plots in Layout

```{r}
(p1 / p2) | p3
```

### Add Auto-Tagging to Composite Plot

```{r}
((p1 / p2) | p3) + 
  plot_annotation(tag_levels = 'I')
```

### Add Inset Plot

```{r}
p3 + inset_element(p2, left = 0.02, bottom = 0.7, right = 0.5, top = 1)
```

### Combine Patchwork with ggthemes

```{r}
patchwork <- (p1 / p2) | p3
patchwork & theme_economist()
```
